Renewal and stability in populations structured by remaining years of life

### 2014-08-29 ThanoRepro/ThanoRepro.tex
added new references, updating more manuscript to account for Jim's new manuscript

\noindent where $\omega$ is the highest age. One
produces in this way a probability density function over future death times
for each chronological age, $a$, and the $d(y|a)$- weighted average of $y$ for
a given $a$ returns the familiar remaining life expectancy column of the lifetable,
$e(a)$. Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
this density is useful for decomposing data classified by chronological age. For
example, we can break down a population pyramid into thanatological age classes.
Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data, assuming the
2010 period mortality schedule remains constant. When looking from the
chronological age perspective, this decomposition reveals thanatological
heterogeneity, but when viewed from the thanatological perspective
(Figure~\ref{fig:USdecomp}) it reveals chronological age heterogeneity and the
overall thanatological age profile, the future death flow of the current
population.
Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010
US data.

\begin{figure}
	\caption{2010 US population structure}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological age-structure (years lived) decomposed by
		thanatological age (years left).}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological age-structure (years left) decomposed by chronological
		age (years lived).}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

If $P_a$ is an chronological age-classified discrete population count, we can
derive $P_y$, a thanatological age-classified population count, as follows:

\begin{align}
\label{eq:transform}
P_y =& \sum_{a=0}^\omega P_a \frac{d_{a+y}+d_{a+y+1}}{L_a+L_{a+1}}
\intertext{and of course}
\sum P_y =& \sum P_a
\end{align}

\citet{brouard1989mouvements}, and later
\citet{vaupel2009life}, showed that in stationary populations the chronological
and thanatological age structures are identical.
It bears repeating that lifetable-derived thanatological age structure is
probabilistic until after death, since it refers to the future according to particular assumptions about mortality. As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking glance at a population attrition given the age-structure and
lifetable of a particular moment, whereas Figure~\ref{fig:USdecomp} is
reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also
migration and to a lesser degree attrition. I explore population structure by remaining years of life
for its own sake in a different paper.\footnote{Poster Session 6 (Friday
mid-day))} Note that any count classified by chronological age can be
reclassified in this way, as long as you have a lifetable that plausibly represents the population whose data you wish to reclassify. This means we can derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
to age-classified birth counts, $B_a$, to get $B_y$ and again to
exposure-to-risk, $E_a$ (here we take exposures from all ages), to get $E_y$ and then dividing:

>>>>

\noindent where $\omega$ is the highest age. One
produces in this way a probability density function over future death times
for each chronological age, $a$, and the $d(y|a)$- weighted average of $y$ for
a given $a$ returns the familiar remaining life expectancy column of the lifetable,
$e(a)$. If $P(a)$ is an chronological age-classified population count,
we can derive $P(y)$, a thanatological age-classified population count, as
follows:

\begin{align}
\label{eq:transform}
P(y) =& \int_0^\infty P(a) d(y | a) \dd a
\end{align}

Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
this density is useful for decomposing data classified by chronological age. For
example, we can break down a population pyramid into thanatological age classes.
Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data, assuming the
2010 period mortality schedule remains constant. When looking from the
chronological age perspective, this decomposition reveals thanatological
heterogeneity, but when viewed from the thanatological perspective
(Figure~\ref{fig:USdecomp}) it reveals chronological age heterogeneity and the
overall thanatological age profile, the future death flow of the current
population.
Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010
US data.

\begin{figure}[ht!]
	\caption{2010 US population structure}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological age-structure (years lived) decomposed by
		thanatological age (years left).}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological age-structure (years left) decomposed by chronological
		age (years lived).}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

\citet{brouard1986structure} applied this method to highlight regional
differences in French age structure, at times combining this projective method
with historical stocks. \citet{brouard1989mouvements}, and later
\citet{vaupel2009life}, showed that in stationary populations the chronological
and thanatological age structures are identical, and this provides a
limited basis for comparison between the two perspectives. \citet{wachter2014essential} hints at, and
\citet{vaupel2014stable} expands on how to make chronological and
thanatological age structure commensurate under a constant growth rate, $r$.

This result provides a standard on which to compare chronological and thanatological age structure. It bears repeating that lifetable-derived thanatological age structure is probabilistic until after death, since it refers to the future according to an assumed mortality pattern.
As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking
glance at population attrition given the population stock and lifetable of a
particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also migration and to a lesser degree attrition. Note that any count classified by chronological age can be
reclassified in this way, as long as you have a lifetable that plausibly represents the population whose data you wish to reclassify. This means we can derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
to age-classified birth counts, $B_a$, to get $B_y$ and again to
exposure-to-risk, $E_a$ (here we take exposures from all ages), to get $E_y$ and then dividing:
